---
title: "sim-mad"
author: "Weigang Qiu"
date: "5/21/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1. Construct MAD populations with and without crossing-over (recombination)

Generate maximally divergent 20-bits long binary sequences using genetic algorithms:

```{r}
library(GA)
library(stringdist)
library(tidyverse)
library(ape)

len.seq <- 20 # length of sequences
num.alle <- 10 # number of alleles
size.pair <- num.alle * (num.alle-1)/2

# Fitness function, to maximize the minimal Hamming distance:
epi.fit <- function(x) { 
  x.list <- split(x, rep(1:num.alle, each=len.seq))
  min(seq_distmatrix(x.list, method = "hamming"))/len.seq
}

# Run genetic algorithm with default parameters:
mad.ga <- ga(type = "binary", fitness = epi.fit, nBits = len.seq * num.alle, maxiter = 1000)

plot(mad.ga)


# Run genetic algorithm without recombination:
mad.clonal <- ga(type = "binary", fitness = epi.fit, nBits = len.seq * num.alle, maxiter = 1000, pcrossover = 0)
plot(mad.clonal)

top.pop = mad.ga@solution[1,]
top.clonal = mad.clonal@solution[1,]

# Split into 20-bit strings & get consensus
x.list <- split(top.pop, rep(1:num.alle, each=len.seq))
mad.alleles <- t(as.data.frame(x.list))

x.clonal <- split(top.clonal, rep(1:num.alle, each=len.seq))
clonal.alleles <- t(as.data.frame(x.clonal))

# Obtain majority-rule consensus sequences
consense <- apply(mad.alleles,2, function(x) {table(x)} %>% which.max()-1)

```

## Part 2. Compare seq separation

Compare among variants with crossing-over, without crossing-over, and randomly permuted (with crossing-over). Recombination creates more optimal separation (higher median with low variance).

```{r}

dist.optimal <- seq_distmatrix(x.list, method = "hamming")/len.seq
dist.clonal <- seq_distmatrix(x.clonal, method = "hamming")/len.seq

mad.shuffled <- apply(mad.alleles, 2, sample) # permute by column
xy.list <- split(mad.shuffled, seq(nrow(mad.shuffled)))
dist.shuffled <- seq_distmatrix(xy.list, method = "hamming")/len.seq

three.dist <- tibble(dist = c(dist.optimal, dist.shuffled, dist.clonal), class = c(rep("mad", size.pair), rep("permuted",size.pair), c(rep("clonal", size.pair))))

two.dist <- three.dist %>% filter(class != "permuted")

med <- two.dist %>% group_by(class) %>% summarise(med = median(dist))

two.dist %>% ggplot(aes(x=dist, fill = class)) + geom_density(alpha=0.5) + theme_bw() + geom_vline(data=med, aes(xintercept = med, color=class), size=2)

two.dist %>% ggplot(aes(x = class, y = dist)) + geom_boxplot() + geom_jitter(shape=1) + theme_bw() + geom_hline(yintercept = 0.5, linetype = 2)

t.test(data = two.dist, dist ~ class)

```

## Part 3. Obtain centroids

Find broadly cross-reactive centroids, again using genetic algorithm:

```{r}

# Fitness function, minimize max distance
center.fit <- function(x) {
  ref.alleles <- split(top.pop, rep(1:num.alle, each=len.seq));
  -1 * max(seq_dist(x, ref.alleles, method = "hamming")/len.seq);
}
center.ga <- ga(type = "binary", fitness = center.fit, nBits = len.seq, popSize = 100, maxiter = 200);

plot(center.ga)
centroid <- center.ga@solution[1,]

```

## Part 4. Validation with tree

```{r}
x.list[[11]] <- consense
x.list[[12]] <- centroid

names(x.list) <- c("A01", "A02", "A03", "A04", "A05", "A06", "A07", "A08", "A09", "A10", "Consense", "Centroid")

# distance matrix
x.dist <- seq_distmatrix(x.list, method = "hamming")
x.dist

# show varints
t(as.data.frame(x.list))

# tree
plot(bionj(x.dist), type = "u", no.margin = T, edge.width = 2)
add.scale.bar()

```

## Part 5. Show session info

```{r}
sessionInfo()
```